### 2016-07-05 text/abstraction_relation.tex
Add more on homomorphisms.
Update with some work-in-progress for abstraction relation.

@@

@@

@@

Let $iom = \big\langle \mathit{NAC}, \mathit{Input}$, $\mathit{Output}$, $\mathit{trace}\big\rangle \in \textsc{IOM}$ be an input-output model, also known as a transformation execution. \bentley{Make sure this terminology is consistent.}

Let also $pc = \big\langle \mathit{Input}$, $\mathit{Output}$, $\mathit{trace}\big\rangle \in \textsc{IOM}  \in
\textsc{PathConds}$ be a path condition.

@@

We have that $iom$ is abstracted by $pc$, noted $ex\Vvdash pc$, if and only if:

NAC doesn't edge match iom's Input
Input edge matches iom's Input

\begin{multline}
\label{eq:abstr_input_output}
\big(\forall rl\in Rules\;.\;Match(rl) \vartriangleleft Input^*\big) \;\land\; Output \blacktriangleleft Apply
\end{multline}
\center{and}
\begin{multline}
\label{eq:abstr_trace}
\big(\forall trc\in Components(pc|_{trace})\;.\;trc\vartriangleleft ex\big)\;\land\;\\
\big(\forall e'\in E'\, \exists e\in E\;.\;\tau'(e')=trace \Longrightarrow \tau(e)=trace\big)
\end{multline} 
% where for all $rl,rl'\in Rules$, if $rl\cong rl'$ then there exist two injective typed graph homomorphisms $f$ and $g$ such that $Match(rl)\stackrel{f}{\vartriangleleft} Input^*$, $Match(rl)\stackrel{g}{\vartriangleleft} Input^*$ and $CoDom(f)\neq CoDom(g)$.
% $/pc/ = Apply_{ex}\cup ex|_{trace}$ is the output part of $pc$, including symbolic traceability links.

>>>>

%\item Path conditions onto transformation executions
%\begin{itemize}
%\item Each rule in graph:
%\begin{itemize}
%\item As in paper. Match component is injectively found. Apply component is surjectively found.
%\item For traceability links, PC links are found injectively. Transformation executions links are found surjectively. Not a bijection because multiple rule executions could be present in transformation execution.
%
%\item Question: How to handle unification of matching. Seems to be straightforward before/after must match.
%\end{itemize}
%\end{itemize}

@@

Finally, we define a function getComponents: PC $\rightarrow$ Set(Typed Graph). This function is used for the abstraction relation, as it returns a set of subsets from the match graph of the path condition. These components are collections of the match graph of rules that have been overlaid on top of each other. These components must be matched as a unit, as their elements must be injectively matched. However, elements from different components may match onto the same target vertex in the abstraction relation. More information on how the components are constructed will be provided in the section on path condition construction. \bentley{This is not elegant at all...}

@@

LOOK AT COMPONENTS IN PC MATCH GRAPH AS ELEMENTS AND TRACE LINKS

THEN LOOK AT OUTPUT ELEMENTS AND TRACE LINKS IN IOM

REDO STRUCTURES TO ONLY BE ONE VERTEX AND EDGE SET, AND HAVE SUBSETS FOR MATCH, APPLY

@@

Let $iom = \big\langle \mathit{Input}$, $\mathit{Output}$, $\mathit{trace}\big\rangle \in \textsc{IOM}$ be an input-output model, also known as a transformation execution. \bentley{Make sure this terminology is consistent.}

Let also $pc = \big\langle  \mathit{NAC}$, $\mathit{Input}$, $\mathit{Output}$, $\mathit{trace}\big\rangle  \in \textsc{PathConds}$ be a path condition.

@@

We have that $iom$ is abstracted by $pc$, noted $ex\Vvdash pc$, if and only if the following conditions are all true:

\noindent\makebox[\linewidth]{\rule{\linewidth}{0.4pt}}
The PC's NAC does not match onto the iom's Input

\begin{multline}
\label{eq:abstr_no_nac}
\nexists \big(\mathit{PC}_{NAC} \stackrel{inj}{\vartriangleleft} \mathit{iom}_{Input}^*\big)
\end{multline}

\noindent\makebox[\linewidth]{\rule{\linewidth}{0.4pt}}

Each component in the PC:
Elements matches on the iom's Input
Traceability links can be found as well

\begin{multline}
\label{eq:abstr_input_output}
\big(\forall \mathit{comp} \in \mathit{getComponents(PC)} \vartriangleleft Input^* : \mathit{comp} \stackrel{inj}{\vartriangleleft} \mathit{iom}_{Input}^*\big)
\end{multline}

\noindent\makebox[\linewidth]{\rule{\linewidth}{0.4pt}}

Elements in the iom's Output graph must surjectively match on the Output graph of the PC
Trace links in iom can be found in PC

\begin{multline}
\label{eq:abstr_apply_surjective_match}
\mathit{iom}_{Output} \stackrel{surj}{\vartriangleleft} \mathit{PC}_{Output}
\end{multline}

\noindent\makebox[\linewidth]{\rule{\linewidth}{0.4pt}}

%\begin{multline}
%\label{eq:abstr_trace}
%
%\forall \mathit{PCTrace} \in \mathit{PC}_{trace} :
%
%\exists \mathit{IOMTrace} | \mathit{IOMTrace}
%
%\big(\forall trc\in Components(pc|_{trace})\;.\;trc\vartriangleleft ex\big)\;\land\;\\
%\big(\forall e'\in E'\, \exists e\in E\;.\;\tau'(e')=trace \Longrightarrow \tau(e)=trace\big)
%\end{multline} 
% where for all $rl,rl'\in Rules$, if $rl\cong rl'$ then there exist two injective typed graph homomorphisms $f$ and $g$ such that $Match(rl)\stackrel{f}{\vartriangleleft} Input^*$, $Match(rl)\stackrel{g}{\vartriangleleft} Input^*$ and $CoDom(f)\neq CoDom(g)$.
% $/pc/ = Apply_{ex}\cup ex|_{trace}$ is the output part of $pc$, including symbolic traceability links.

### 2016-07-05 text/formal_background.tex
Update with some work-in-progress for abstraction relation.

Note that the our verification technique is geared toward model-driven engineering. Therefore, the types of vertices and edges in our typed graphs will be drawn from a particular \textit{metamodel}.

Sample metamodels for our running example can be seen in figures above \bentley{Fix.}. Note that DSLTrans rules (to be formally defined) will combine typed graphs with potentially different metamodels.

We assume the existence of a number of functions in our formalization to aid the treatment of typing:

@@

\begin{itemize}
\item The purpose of this function is to assist in handling polymorphism in the metamodel
\item $\mathit{matchesOver}(T, T) \rightarrow \mathit{true}$
\item $\mathit{matchesOver}(T, T') \rightarrow \mathit{true}$ iff T' is a subclass of T in some defined partial ordering $\leq$ given by the metamodel
\item Otherwise, $\mathit{false}$
\end{itemize}
\end{itemize}

@@

\subsubsection*{Typed Subgraph}
We now define the useful notion of typed subgraph. As expected, a typed subgraph is simply a restriction of a typed graph to some of its vertices and edges. \bentley{Is this needed?}

@@

Our technique also requires a form of graph homomorphism that primarily focuses on matching the edges of the graphs. This homomorphism will be an injective\footnote{For a terminology refresh, \textit{injective matching} means there is a one-to-one correspondence between the source and target vertices. \textit{Surjective matching} means that all target vertices must be matched by at least one source vertex. }match regarding the edges, but note that a particular vertex in the source graph may match onto multiple vertices in the target graph. \bentley{What's the term for this?}

@@

Given a graph, its transitive closure includes, besides the original graph, all the edges belonging to the transitive closure of indirect links in that graph. Note that these transitive edges are given by the function $\mathit{isIndirect}$.

@@

>>>>

Note that the our verification technique is geared toward model-driven engineering. Therefore, the types of vertices and edges in our typed graphs will be drawn from a particular \textit{metamodel}. Sample metamodels for our running example can be seen in figures above \bentley{Fix.}. Note that DSLTrans rules (to be formally defined) will combine typed graphs with potentially different metamodels.

We assume that this metamodel provides the necessary vertex and edge types, with any collisions resolved by adding the metamodel name. As well, we assume the presence of a partial ordering $\leq$ on the vertex and edge types for subtyping information.

We require specially typed edges for some DSLTrans constructs. Thus, we include the edge types \{\textit{trace}, and \textit{backward}\} in ET, resolving collisions as needed. These edge types will be explained as needed. We also insert the ordering that \textit{trace} is a subtype of \textit{backward}.

For convenience, we define some utility functions in our formalization to aid the treatment of typing:

@@

\begin{itemize}
\item The purpose of this function is to assist in handling polymorphism in the metamodel, and to handle matching for the special DSLTrans types
\item $\mathit{matchesOver}(T, T) \rightarrow \mathit{true}$
\item $\mathit{matchesOver}(T, T') \rightarrow \mathit{true}$ iff T' is a subclass of T in some defined partial ordering $\leq$ given by the metamodel
\item Otherwise, $\mathit{false}$
\end{itemize}
\end{itemize}

@@

\subsubsection*{Typed Subgraph}
We now define the useful notion of typed subgraph. As expected, a typed subgraph is simply a restriction of a typed graph to some of its vertices and edges. This will be needed to partition DSLTrans constructs such as rules into logical components.

@@

Our technique also requires a form of graph homomorphism that primarily focuses on matching the edges of the graphs. This homomorphism will be an injective\footnote{For a terminology refresh, \textit{injective matching} means there is a one-to-one correspondence between the source and target vertices. \textit{Surjective matching} means that all target vertices must be matched by at least one source vertex.} match regarding the edges, but note that a particular vertex in the source graph may match onto multiple vertices in the target graph. \bentley{What's the term for this?}

@@

Given a graph, its transitive closure includes, besides the original graph, all the edges belonging to the transitive closure of indirect links in that graph. Note that these transitive edges are identified by the function $\mathit{isIndirect}$.

@@

\subsubsection{Link Homomorphism}

\begin{itemize}
\item Backward links injectively match from contract into PC onto traceability links. Source and target of backward links on contract must match the source and target of the traceability link of the PC.
\end{itemize}

We must also define a function to match the traceability and backward link structures.

\begin{definition}{Link Homomorphism\\}
\label{def:link_homomorphism}
Consider a set of links $L = \{E_{L}, (s_{L}, t_{L})\}$. Note that these could be backward links or traceability links.

We define a function $f$ to match $L$ over $T$, such that:

\begin{itemize}
\item $\forall \mathit{sourceLink} \in L:$
\begin{itemize}
\item $\exists \mathit{targetLink} \in T | f(s(\mathit{sourceLink})) = s(targetLink), $
\end{itemize}
\end{itemize} 
\end{definition}
